---
title: "Assignment 11 Approach"
format: html
editor: visual
---

## Assignment 11

In assignment 3A, I created a simple movie recommender system. I've iterated on this system, creating a content-based filtering system that recommends similar movies with high average ratings.

I augmented the data by adding genre information about each movie. I put together my own algorithm, then used recommenderlab. Finally, I compare the recommendations made with an actual model to those using simple math.

```{r, warning = FALSE}
library(tidyverse)
library(recommenderlab)
```

Reading the original data set:

```{r}

sql_url <- "https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/moviedata_CSV.csv"  
#read and convert blanks and NULL to NA  
movie_data <- read.csv(sql_url, na = c("", "NA", "null", "NULL")) 
```

### Cleaning (reproduced from project 3A)

```{r}
#converting to tibble  
movie_data <- as_tibble(movie_data)   

#pivot longer
long_movie_data <- movie_data |> pivot_longer(   -viewer,   names_to = "movie",   values_to = "rating" )  


#And add the average for each movie 

new_ratings <- long_movie_data |> 
    mutate(avg_rating_global = round(mean(rating, na.rm = TRUE), 1)) |>
    group_by(movie) |> 
    mutate(avg_rating_movie = round(mean(rating, na.rm = TRUE), 1)) |> 
    ungroup()


new_ratings <- long_movie_data |> 
    mutate(avg_rating_global = round(mean(rating, na.rm = TRUE), 1)) |>
    group_by(movie) |>
    mutate(avg_rating_movie = round(mean(rating, na.rm = TRUE), 1)) |> 
    ungroup() |> 
    group_by(viewer) |>
    mutate(viewer_avg = round(mean(rating, na.rm = TRUE), 1)) |> 
    ungroup() |>
    mutate(user_diff = viewer_avg - avg_rating_global)

#predicted ratings

new_ratings <- new_ratings |> mutate(predicted_rating = (ifelse(is.na(rating), user_diff + avg_rating_movie, NA)))
```

### Creating a key

Adding an incredibly nuanced genre category to match with ratings list.

```{r}
movie_list <- long_movie_data |> 
  distinct(movie) |> mutate(genre = "holiday")

movie_list[1:3, 2] <- "marvel"

movie_list[4:6, 2] <- "dc"

```

### Joining

Adding a genre to the original data frame:

```{r}
includes_genres <- left_join(new_ratings, movie_list, by = "movie") 

genre_average <- includes_genres |>
  group_by(viewer, genre) |>
  mutate(avg_genre = na_if(round(mean(rating, na.rm = TRUE), 1), NaN)) |> 
  ungroup()
  

#creating an index of favorite genres
genre_average_only <- includes_genres |>
  group_by(viewer, genre) |>
  summarise(gen_avg = (round(mean(rating, na.rm = TRUE), 1))) |>
  arrange(viewer, desc(gen_avg))

genre_average_only
  
```

### Sorting by predicted rating and preference

Find the users' genre order of preference and recommend the next movie in that genre they should watch based on genre and predicted rating.

Selecting the relevant columns:

```{r}
predictions_list <- genre_average |>
  select(viewer, movie, rating, predicted_rating, genre, avg_genre) |>
  arrange(viewer, desc(avg_genre), desc(predicted_rating))
```

### Incorporating genre (items) into predictions

Because these are arranged by genre preference, we can just choose the first entry for each user where the rating is NA.

```{r}
genre_predictions <- predictions_list |>
  filter(is.na(rating)) |>
  group_by(viewer) |>
  filter(row_number()==1)

genre_predictions
```

The result is a simple content-based rating system. It's similar to project 3A, but it predicts the highest-rated movie of the genre the user is most interested in. I could have gone by average movie rating for that movie instead of predicted rating (the result would have been the same) to make the system purely item-based.

## Using the recommenderlab package

After several turns with Claude, here is a version of the recommenderlab package that actually functions (I hesitate to use the word "works"). I will attempt to explain/understand what's happening in the code below

Step 1: This turns the rating into a double (not necessary) and fills in one value for "Best rating."

Step 2: Creates a matrix of full ratings for each user and pivots wider. The distinct function also isn't necessary, but may be for a larger or messier data set.

```{r}

# ── 1. Coalesce actual + predicted rating ──────────────────────────────────────

predictions_list <- predictions_list |>
  mutate(best_rating = coalesce(as.double(rating), predicted_rating))

# ── 2. Build rating matrix ─────────────────────────────────────────────────────

rating_matrix <- predictions_list |>
  select(viewer, movie, best_rating) |>
  distinct(viewer, movie, .keep_all = TRUE) |>
  pivot_wider(names_from = movie, values_from = best_rating) |>
  column_to_rownames("viewer") |>
  as.matrix()



```

Step 3: Cosine is a common way to measure similarity with recommender systems

```{r}
# ── 3. Genre similarity matrix ─────────────────────────────────────────────────

cosine_sim <- function(mat) {
  sim   <- mat %*% t(mat)
  norms <- sqrt(rowSums(mat^2))
  sim / outer(norms, norms)
}

genre_matrix <- predictions_list |>
  distinct(movie, genre) |>
  mutate(value = 1) |>
  pivot_wider(names_from = genre, values_from = value, values_fill = 0) |>
  column_to_rownames("movie") |>
  as.matrix()

#A wide DF where 1 denotes whether something belongs to a genre or not
genre_sim <- cosine_sim(genre_matrix)

```

Step 4 creates a rating matrix that shows how correlated different ratings are. For example, superman and birds of prey have a -.9364 correlation, based on users who rated both movies. Someone who rated Superman well might rate Birds of Prey poorly.

```{r}
# ── 4. Rating similarity matrix ────────────────────────────────────────────────

rating_sim <- cor(rating_matrix, use = "pairwise.complete.obs")
rating_sim[is.na(rating_sim)] <- 0

```

Step 5 finds movies that exist in both genre and rating matrix (in theory, there could be unrated movies that have a genre listed, though that's not true for this data)

Blended_sim weights how important rating and genre are in defining items for the rating system. The original split was .7/.3, but I changed it to .4/.6, with a heavier weight toward genre.

```{r}
# ── 5. Blend similarities 70/30 ────────────────────────────────────────────────
shared_movies <- intersect(rownames(genre_sim), colnames(rating_matrix))

blended_sim <- 0.4 * rating_sim[shared_movies, shared_movies] +
               0.6 * genre_sim[shared_movies, shared_movies]
```

Step 6 creates the IBCF recommendation.

-   First two functions: pulls all the viewers' ratings (actual and predicted); then pulls only the actual ratings

-   Scores - scores movies the user hasn't actually seen.

-   Sims - Finds similarities between target movie and all other movies.

-   Rated mask is a boolean function that tells whether the user has rated a single movie (otherwise, predictions are not possible)

-   Denom is the weighted avarage formula - it weight similarity to other movies, and attempts to figure out how a user would related a movie based on how they've rated other movies

```{r}
# ── 6. IBCF scoring function — bypasses recommenderlab's broken list export ────
predict_ibcf <- function(viewer_name, rating_mat, sim_mat, n = 10) {
  
  viewer_ratings <- rating_mat[viewer_name, shared_movies]
  
  # Original (non-coalesced) ratings — NA = truly unrated
  original_ratings <- predictions_list |>
    filter(viewer == viewer_name) |>
    select(movie, rating) |>
    deframe()
  
  # Score every movie for this viewer
  scores <- map_dbl(shared_movies, function(target_movie) {
    
    # Only recommend movies the viewer actually hasn't rated
    if (!is.na(original_ratings[target_movie])) return(NA_real_)
    
    # Similarities between target movie and all other movies
    sims <- sim_mat[target_movie, shared_movies]
    
    # Only use movies this viewer HAS rated as neighbors
    rated_mask <- !is.na(viewer_ratings)
    if (sum(rated_mask) == 0) return(NA_real_)
    
    neighbor_sims    <- sims[rated_mask]
    neighbor_ratings <- viewer_ratings[rated_mask]
    
    # Weighted average: high-similarity movies pull harder on the prediction
    denom <- sum(abs(neighbor_sims))
    if (denom == 0) return(NA_real_)
    sum(neighbor_sims * neighbor_ratings) / denom
  })
  
  names(scores) <- shared_movies
  
  # Return top N unrated movies sorted by predicted score
  scores |>
    na.omit() |>
    sort(decreasing = TRUE) |>
    head(n) |>
    enframe(name = "movie", value = "ibcf_score") |>
    mutate(viewer = viewer_name)
}

```

all_viewers - list of all viewers taken from the row names of the rating matrix

recs_df is the predicted ratings of unseen movies based on the algorithm

eval_df tests the model - it *is* predicting against the original set of predicted ratings, which seems incorrect: it should test predictions against the viewer's actual scores.

```{r}
# ── 7. Generate recommendations for all viewers ────────────────────────────────
all_viewers <- rownames(rating_matrix)

recs_df <- map_dfr(all_viewers, ~predict_ibcf(
  viewer_name = .x,
  rating_mat  = rating_matrix,
  sim_mat     = blended_sim,
  n           = 10
)) |>
  select(viewer, movie, ibcf_score)

cat("=== Raw recommendations ===\n")
print(recs_df)

# ── 8. Evaluate: IBCF score vs predicted_rating on originally-unrated cells ────
eval_df <- recs_df |>
  left_join(
    predictions_list |> select(viewer, movie, predicted_rating),
    by = c("viewer", "movie")
  ) |>
  filter(!is.na(predicted_rating))  # only rows where we have ground truth

rmse <- sqrt(mean((eval_df$ibcf_score - eval_df$predicted_rating)^2))
mae  <-      mean( abs(eval_df$ibcf_score - eval_df$predicted_rating))

cat("\nRMSE:", round(rmse, 3), "\n")
cat("MAE: ", round(mae,  3), "\n")
print(eval_df)

# ── 9. Join full context back in ───────────────────────────────────────────────
final_recommendations <- recs_df |>
  left_join(
    predictions_list |>
      distinct(viewer, movie, .keep_all = TRUE) |>
      select(viewer, movie, predicted_rating, genre, avg_genre),
    by = c("viewer", "movie")
  ) |>
  arrange(viewer, desc(avg_genre), desc(ibcf_score))

print(final_recommendations)

rmse
mae
```

The RMSE is really high for a 5-point rating system. This is expected – it's a really small data set. Let's check against the actual ratings.

```{r}
all_viewers <- rownames(rating_matrix)

recs_df <- map_dfr(all_viewers, ~predict_ibcf(
  viewer_name = .x,
  rating_mat  = rating_matrix,
  sim_mat     = blended_sim,
  n           = 10
)) |>
  select(viewer, movie, ibcf_score)

cat("=== Raw recommendations ===\n")
print(recs_df)

# ── 8. Evaluate: IBCF score vs predicted_rating on originally-unrated cells ────
eval_df <- recs_df |>
  left_join(
    predictions_list |> select(viewer, movie, predicted_rating),
    by = c("viewer", "movie")
  ) |>
  filter(!is.na(predicted_rating))  # only rows where we have ground truth

rmse <- sqrt(mean((eval_df$ibcf_score - eval_df$predicted_rating)^2))
mae  <-      mean( abs(eval_df$ibcf_score - eval_df$predicted_rating))
```

The RMSE is similar than the one compared to my predictions.

### Comparing results

```{r}
library(gt)
#the model


actual_final_recommendations <- final_recommendations |>
  mutate(ibcf_score = round(ibcf_score, 2))|>
  group_by(viewer) |>
  filter(row_number()==1)

actual_final_recommendations |>
  select(viewer, movie, ibcf_score) |>
  as_tibble() |> 
  gt() |>   
  cols_label(
    viewer = "Viewer",
    movie = "Movie",
    ibcf_score = "Predicted Rating"
  ) |>
  tab_header(title = "Predictions Using Recommenderlab")

#my simple data frame
genre_predictions |>
  select(viewer, movie, predicted_rating) |>
   as_tibble() |> 
  gt(rowname_col = NULL) |>   
  cols_label( 
    viewer = "Viewer",
    movie = "Movie",
    predicted_rating = "Predicted Rating") |>
  tab_header(title = "Predictions Using Math")
```

Overall, the predicted ratings are somewhat different, probably due to the weighted averages. All the recommendations are the same except for Jeff's - the algorithm recommends *The Holiday* instead of *The Grinch*, which is kind of puzzling because they are both holiday movies.

Claude Sonnet 4.6. (2026). \[Large language model\]. https://gemini.google.com. Accessed April 23 & 24, 2026.
