---
title: "Project 7 approach"
author: "Sam Barbaro"
format: html
editor: visual
---

## Approach

I've created HTML and JSON files and uploaded them to GitHub. These files feature information about books I edited. I will load them here and compare them to see if they are identical and, if not, where they differ.

books.html: <https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/books.html>

books.json: <https://github.com/samanthabarbaro/data607/blob/main/books.json>

```{r}
library(tidyverse)
library(rvest)
library(jsonlite)

books_h <- read_html("https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/books.html")


#turn into a data frame
html_df <- books_h |>
  html_table(fill = TRUE) %>%
  .[[1]] 


#Is this a data frame?
is_tibble(html_df)

#load json file and turn it into a data frame
books_j <- fromJSON("https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/books.json")

```

## Testing whether the data are identical

These shouldn't be 100% identical since I used a semicolon in the HTML file, but not the JSON file.

```{r}
#are they identical? 
identical_data <- all.equal(books_j, html_df)

#not exactly, but they're close

identical_data

#A true or false function to check whether they're identical
identical(books_j, html_df)

#They are not identical

#Let's make them more identical

html_df_2 <- html_df |>
    separate(col = Authors, into = c("Author_1", "Author_2"), sep = ";") 

#I really should have left the semicolon as a separator

books_j2 <- books_j |>
    separate(col = Authors, into = c("Author_1", "Author_2"), sep = ",") 

#replacing individual values

books_j2[2, 2] <- "Angela Wood, Ph.D"

books_j2[2, 3] <- NA

#fixing some of the characters, which converted strangely
html_df_2[1, 1] <- "Threshold Concepts in Women’s and Gender Studies"
books_j2[1, 1] <- "Threshold Concepts in Women’s and Gender Studies"


identical_data_2 <- all.equal(books_j2, html_df_2)

identical_data_2

diffs <- setdiff(books_j2, html_df_2)
diffs

#The tables have fewer differences, but still aren't exactly the same

identical(books_j2, html_df_2)
```

Google Gemini. (2026). Gemini 3 Flash \[Large language model\].\
https://gemini.google.com. Accessed March 12, 2026.
