---
title: "Assignment 11 Approach"
format: html
editor: visual
---

## Assignment 11

In assignment 3A, I created a simple movie recommender system. I will iterate on this system, creating a content-based filtering system that recommends similar movies with high average ratings.

I will augment the data (for example, by adding more information about each items and adding more movies). I will either put together my own algorithm or use a system like the recommenderlab package loaded below.

**One proposed way to do this:**

Create a new table with movies and category. Cross-reference this with the user's highest-rated movie category and recommend the highest-rated movie in that category that the user has not seen. If the user has seen all movies in that category, move on to the next-highest-rated, and so on.

```{r, warning = FALSE}
library(tidyverse)
library(recommenderlab)
```

Reading the original data set:

```{r}

sql_url <- "https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/moviedata_CSV.csv"  
#read and convert blanks and NULL to NA  
movie_data <- read.csv(sql_url, na = c("", "NA", "null", "NULL")) 
```
