---
title: "Assignment 10A"
format: html
editor: visual
---

## Assignment 10 A

We've reproduced the example code for sentiment analysis from chapter two of *Text Mining With R*. Using the sotu library (filled with state of the union speeches up until 2020), we will break up the text and perform a similar sentiment analysis with the speeches of Abraham Lincoln, Franklin D. Roosevelt, and Barack Obama.

```{r}
library(tidyverse)
library(janeaustenr)
library(tidytext)
library(textdata)


tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)



#get_sentiments("afinn")
#get_sentiments("bing")
#get_sentiments("nrc")
```

```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```         
http://saifmohammad.com/WebDocs/Lexicons/NRC-Emotion-Lexicon.zip' Content type 'application/zip' length 25878449 bytes (24.7 MB)
```

```{r}
library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)



ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice

```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)

get_sentiments("bing") %>% 
  count(sentiment)
```

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```

```{r}
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

Silge, J., & Robinson, D. (2017). *Text mining with R: A tidy approach*. O'Reilly Media.

The sotu package contains state of the union addresses.

```{r}
library(sotu)

data(sotu_text)

#make a tidy data frame
sotu_df <- tibble(
  president = sotu_meta$president,
  year = sotu_meta$year,
  text = sotu_text
)

#unnest
tidy_sotu <- sotu_df %>%
  unnest_tokens(word, text)
```

There are hundreds of state of the union addresses. For this exercise, we'll look at three presidents: Abraham Lincoln, Franklin D. Roosevelt, and Barack Obama.

```{r}
obama <- sotu_df |>
  filter (president == "Barack Obama")

d_roosevelt <- sotu_df |>
  filter (president == "Franklin D. Roosevelt")

lincoln <- sotu_df |>
  filter (president == "Abraham Lincoln")
```
