---
title: "Project 3"
author: "Sam Barbaro"
format: html
editor: visual
---

## Project 3 Approach

I will take three untidy, wide data sets shared by my classmates, including theaverage precipitation by city:

<https://en.wikipedia.org/wiki/List_of_cities_by_average_precipitation>

Disney parks monthly attendance, and field goal career stats from NFL statistics <https://www.kaggle.com/datasets/kendallgillies/nflstatistics> (The third one may be too clean to begin wiht and I may switch it out)

For each data set, I will:

-   Reshape data from wide to tidy (long) format, using pivot longer

-   Normalize the variable structure (e.g., all numeric, same number of decimals)

-   Rename variables to follow a consistent naming convention

-   Address missing or inconsistent values

-   Document how I handle missing values

    ### Planned analysis

For **cities by average precipitation**, I will compare seasonal precipitation patterns across cities in the same country, and to find the rainiest and driest cities overall. I will also compare average rainfall across regions. I will also add country via join.

For the **Disney Parks data** I willl look at monthly attendance to identify peak/low travel seasons, compare FL and CA attendance, and calculate YTD totals for each park. I will also visualize monthly attendance across parks. I have added additional parks (Animal Kingdom and Calinfornia Adventure) to compare all CA parks to all FL parks.

With the football data set, I will look at cumulative player stats, ratios, and the relationship of certain stats (such as player height and weight) relate to career length.

```{r}
library(tidyverse)
library(lubridate)
```

## Load the data sets

```{r}
#disney parks
disney_parks <- 
  read.csv("https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/disney_parks_monthly_attendance.csv", 
                          na.strings = c("", "NA", "null", "NULL"), 
                          header = TRUE)


#NFL field goals
field_goal <-  read.csv("https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/Career_Stats_Field_Goal_Kickers.csv", 
                          na.strings = c("", "--", "NA", "null", "NULL"), 
                          header = TRUE)

rainfall <- read.csv("https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/rainfallbycity.csv", na.strings = "", header = TRUE)
```
