---
title: "Project 3 A"
author: "Samantha Barbaro"
format: html
editor: visual
---

## Approach

I will use my movie data set from the last project to create a recommendation system. To do this, I will:

-   find the average rating for all movies

-   Find out how the average of every individual movie compares to the overall average

-   Find out how a user's average rating compares to the overall average

-   Using these numbers, predict ratings for movies users haven't seen

-   Provide a recommendatin by recommending the movie they are likely to rate the highest

Data set:

<https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/moviedata_CSV.csv>

## Loading the data set

```{r}
library(tidyverse)  

sql_url <- "https://raw.githubusercontent.com/samanthabarbaro/data607/refs/heads/main/moviedata_CSV.csv"  
#read and convert blanks and NULL to NA  
movie_data <- read.csv(sql_url, na = c("", "NA", "null", "NULL")) 


```

## Converting to a tibble

```{r}
#converting to tibble  
movie_data <- as_tibble(movie_data)   

print(movie_data) 
```

### Pivot longer

```{r}
#this may be required to analyze the data (making the data into a 2-column format)  

long_movie_data <- movie_data |> pivot_longer(   -viewer,   names_to = "movie",   values_to = "rating" )  

#get some descriptive statistics 
summary(long_movie_data) 
```

## 
