---
title: "Project 4"
format: html
editor: visual
---

## Project 4: Approach

I'm going to clean these spam and ham documents (stripping out stop words, punctuation, etc.) so I can analyze common words. I'll use tools like word cloud and sentiment analysis, and I'll at the length of emails to see if ham or spam is more commonly shorter.

I will then use either tidytext or quanteda to train/test a model and analyze the numbers of true/false positives and negatives. Chapter 6, "Topic Modeling," from [Text Mining with R](https://www.tidytextmining.com/topicmodeling) may be helpful here.

Reading the files from my local drive:

```{r, warning=FALSE}
library(tidyverse)
library(readtext)
library(tidytext)
library(textdata)
library(wordcloud)
library(tidyr)
library(reshape2)

#load the spam files from my local drive
spam_df <- readtext("C:/Users/Sam Barbaro/Desktop/607 - data warehousing/spamham/20021010_spam/spam")

spam_df <- spam_df |>
  mutate(label = "spam")

```

```{r, warning = FALSE}
#loading the ham files

ham_df <- readtext("C:/Users/Sam Barbaro/Desktop/607 - data warehousing/spamham/20030228_easy_ham/easy_ham")

ham_df <- ham_df |>
  mutate(label = "ham")
```
