---
title: "Project 4"
format: html
editor: visual
---

## Project 4

Below, I load, clean, and tokenize spam and ham corpuses. I use word clouds, LDA, and histograms to understand typical document length and language. I then use the Quanteda package to create a prediction model using the Naive Bayes classifier.

Reading the files from my local drive:

```{r, warning=FALSE}
library(tidyverse)
library(readtext)
library(tidytext)
library(textdata)
library(wordcloud)
library(tidyr)
library(reshape2)

#load the spam files from my local drive
spam_df <- readtext("C:/Users/Sam Barbaro/Desktop/607 - data warehousing/spamham/20021010_spam/spam")

spam_df <- spam_df |>
  mutate(label = "spam")

spam <- spam_df |>
  select(text, label) |>
  mutate(email_id = row_number())
```

```{r, warning = FALSE}
#loading the ham files

ham_df <- readtext("C:/Users/Sam Barbaro/Desktop/607 - data warehousing/spamham/20030228_easy_ham/easy_ham")

ham_df <- ham_df |>
  mutate(label = "ham")

ham <- ham_df |>
  select(text, label) |>
  mutate(email_id = row_number())
```

### Unnesting the text

```{r}
tidy_spam <- spam %>%
  group_by(email_id) %>%
  mutate(
    linenumber = row_number(),
    email_id = cur_group_id() 
  ) %>%
  ungroup() %>%
    unnest_tokens(word, text)

tidy_ham <- ham %>%
  group_by(email_id) %>%
  mutate(
    linenumber = row_number(),
    email_id = cur_group_id() 
  ) %>%
  ungroup() %>%
    unnest_tokens(word, text)
```

### Word cloud

What are some of the most common words in spam and ham emails?

```{r}
tidy_spam %>%
  anti_join(stop_words, by = join_by(word)) %>%
  count(word) %>%
  with(wordcloud::wordcloud(word, n, max.words = 100))


tidy_ham %>%
  anti_join(stop_words, by = join_by(word)) %>%
  count(word) %>%
  with(wordcloud::wordcloud(word, n, max.words = 100))
```

This does suggest that I need to get rid of a few more common words beyond stop words (like "from" and "zzz").

### Join dataframes 

```{r}
library(topicmodels)
library(tm)

#combine data frames and remove words that are just numbers
#which occupied the first ~5,000 rows the first time I did this 
#also removing stop words
combined_tidy <- rbind(tidy_ham, tidy_spam) %>%
  anti_join(stop_words, by = join_by(word)) %>%
  filter(!str_detect(word, "[0-9]"))

    
tidy_counts <- combined_tidy %>%
  count(email_id, word)

dtm <- tidy_counts %>%
  cast_dtm(email_id, word, n)

# Run LDA to create a two-topic LDA model (span and ham)
spam_lda <- LDA(dtm, k = 2, control = list(seed = 1234))


```

### Word-topic probabilities

```{r}
spam_topics <- tidy(spam_lda, matrix = "beta")
spam_topics
```

### Graphing the top terms by topic

#2 is ham and 1 is spam

```{r}
spam_top_terms <- spam_topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>% 
  ungroup() %>%
  arrange(topic, -beta)

spam_top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```

This doesn't give us a ton more than the word cloud did.

### Analyzing email length

```{r}
set.seed(100)

#taking a sample because this takes a long time
unique_ids <- unique(tidy_spam$email_id)

sampled_ids <- sample(unique_ids, 0.4 * length(unique_ids))

#plotting histograms

tidy_spam |> filter(email_id %in% sampled_ids) |>
  count(email_id, name = "total_words") |>
  ggplot(aes(x = total_words)) +
  geom_histogram() + 
  labs(Title = "Spam Email Length Frequency")

unique_ids <- unique(tidy_ham$email_id)

sampled_ids <- sample(unique_ids, 0.4 * length(unique_ids))

tidy_ham |> filter(email_id %in% sampled_ids) |>
  count(email_id, name = "total_words") |>
  ggplot(aes(x = total_words)) +
  geom_histogram() +
  labs(Title = "Ham Email Length Frequency")



#mean word counts
spam_counts <- tidy_spam |>
  count(email_id, name = "total_words") 

spam_counts |> summarise(mean(total_words))

ham_counts <- tidy_ham |>
  count(email_id, name = "total_words") 

ham_counts |> summarise(mean(total_words))
```

On average, Spam emails tend to be longer.

Let's build a model.

### Quanteda Naive Bayes classifier

[This Quanteda tutorial](https://tutorials.quanteda.io/machine-learning/nb/) uses movie review data in a format similar to the spam/ham corpus. I've modified it below to only use a portion of the data.

```{r}
library(caret)
library(quanteda)
library(quanteda.textmodels)

#data prep
#combined spam_df and ham_df (these have the full text in a single cell)
#create a new email id so everything has its own id once tokenized

spam_ham <- rbind(spam_df, ham_df)

spam_ham <- spam_ham |>
  select(text, label) |>
  mutate(email_id = row_number())

#format as corpus (part of the quanteda package)
spam_corpus <- corpus(spam_ham, text_field = "text")

#this data set is huge, so we'll use about 10% to test 
#generate 300 numbers without replacement (10% of the data)
set.seed(300)
id_train <- sample(1:3000, 300, replace = FALSE)
head(id_train, 10)

#separate set of 100 numbers for test set, so we don't test on training data
remaining_ids <- setdiff(1:3000, id_train)


id_test <- sample(remaining_ids, 100, replace = FALSE)


spam_corpus$email_id <- 1:ndoc(spam_corpus)

# tokenize texts
toks_spam <- tokens(spam_corpus, remove_punct = TRUE, remove_number = TRUE) %>% 
               tokens_remove(pattern = stopwords("en")) %>% 
               tokens_wordstem()

dfmt_spam <- dfm(toks_spam)

# get training set
spam_training <- dfm_subset(dfmt_spam, email_id %in% id_train)

# get test set 
spam_test <- dfm_subset(dfmt_spam, email_id %in% id_test)

```

Train the model:

```{r}


tmod_nb <- textmodel_nb(spam_training, spam_training$label)

summary(tmod_nb)
```

```{r}
#test the model
dfmat_matched <- dfm_match(spam_test, features = featnames(spam_training))

#inspect how well the classification worked.
#Also, the point at which I stopped changing the labels from the sample code

actual_class <- dfmat_matched$label
predicted_class <- predict(tmod_nb, newdata = dfmat_matched)

#the mini confusion matrix
tab_class <- table(actual_class, predicted_class)
tab_class
```

Confusion matrix (positive class is ham):

```{r}
confusionMatrix(tab_class)

confusionMatrix(tab_class, mode = "everything")

```

90% accuracy, 79 true positives, 8 false negatives, 2 false positives, and 11 true negatives. As a spam filter, 8 non-spam emails are going into the spam folder and 2 spam messages sneak through. No one wants 9 legitimate emails out of 87 to go in their spam filter. I wonder if the results woud be better if the sample size were larger.

In retrospect, I could have taken other steps to prep the data. I didn't filter out numbers, short words (though these are filtered by stop words), or nonsense strings, which appear frequently in older email messages.

### Sources

Silge, J., & Robinson, D. (2017). *Text mining with R: A tidy approach*. O'Reilly Media.

Watanabe, K & Müller, S. *Quanteda Tutorials.* <https://tutorials.quanteda.io/machine-learning/nb/>.

Google Gemini. (2026). Gemini 3 Flash \[Large language model\].https://gemini.google.com. Accessed May 1, 2026.
